---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1: The distribution of expenses is very lopsided with about 95% of the expenses being $5,000 or less. I would make the increments maybe by 100 in the beginning instead of going straight to 500 (like add in 100-200, 200-300,etc) and then maybe by thousands instead of going straight to 5000 **

```{r}
maryland_expenses_with_chunks <- maryland_expenses %>% 
  mutate(
    expense_chunk=case_when(
      amount <= 0 ~ "negative", 
      amount <= 500  ~ "0-500", 
      amount <= 5000  ~ "500-5000",
      amount <= 50000  ~ "5001-50000",
      amount <= 100000 ~ "50001-100000",
      amount <= 500000 ~ "100001-500000"
    )
)
```
``` {r}
maryland_expenses_with_chunks %>%
  group_by(expense_chunk) %>%
  summarise(
    count_expense_chunk = n() 
  )
``` 
**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2: Atmospheric and Oceanic Science offers the most courses with a total of 66 being offered by them** 

```{r}
umd_courses %>% 
  filter(str_detect(title, "Climate")) %>%
  group_by(department) %>%
  summarise( 
    count_department = n() ) %>%
  arrange(desc(count_department))
```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3: Economics having 4 classes on climate is pretty interesting because I'd like to learn more about the connection between the climate and the economy and how the climate affects businesses across the world** 

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.October as well as May and June are the highest expense months which are months that occur right before an election (As they are held in July and November)**

```{r}
 maryland_expenses_with_date <- maryland_expenses_with_state %>% 
  mutate(expenditure_date=mdy(expenditure_date))
```
```{r}
maryland_expenses_with_date %>%
  filter(committee_name == "Moore  Wes For Maryland") %>% 
  mutate(month = floor_date(expenditure_date, "month")) %>%
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount)) 
```

**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5: Which date had the highest amount of expenditures? The answer is January 11th 2022 with just short of $3 million on that day (2,839,313.10) ** 
```{r}
maryland_expenses_with_date %>%
  group_by(expenditure_date) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```
