---
title: "lab_12"
author: "derek willis"
date: "2023-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library (rvest)
```

## You will need

* tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**A1.The leading word in the answer is failing (at approximately 4.6%).Broadly the top 10 words describe some of the most common language used in addressing misconduct in a legal proceeding. ** 

```{r}
md_attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")
unique_words <- md_attorney_sanctions %>% select(text) %>%
  unnest_tokens(word, text)
View(unique_words)

unique_words %>%
  anti_join(stop_words) %>%
  group_by(word) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**A2.The most common trigram is "attorney trust account" which appears approximately 5.8% of the time. In legal terms the phrase means "a type of bank account that recieves funds from an attorney in a law firm on behalf of a client or third person" per the MD courts online definition.  ** 

```{r}
md_attorney_sanctions %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3) %>%
  separate(trigram, c("word1", "word2", "word3"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word) %>%
  mutate(trigram = paste(word1, word2, word3, sep=" ")) %>%
  group_by(trigram) %>%
  tally(sort=TRUE) %>%
  mutate(percent = (n/sum(n))*100) %>%
  top_n(10)
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**A3. The produced data suggests that there were more breaches of the attorney trust account within 2021 as compared to other years. From 2019 to 2021 the number of times that the phrase appeared increasingly went up each year to reach the peak in 2021 before going back down a little in 2022 and skyrocketing down in 2023. Essentially this means that there was a trend established from 2019 to 2021 of more and more attorneys getting caught mishandling their clients money and not responsibly handling the situation as they are expected to. For example, David Caplan was disbarred in 2022 for keeping fees and not using them for their necessay payments and misrepresenting himself. To clarify the most important reasons attorneys get sanctioned, I would definitely try to look into what phrases come after the word "for" in the data set. Most times the format is stated as "x was disbarred for x".** 

```{r}
md_attorney_sanctions %>%
  filter (str_detect(text, "attorney trust account" )) %>%
  group_by(fiscal_year)%>%
  summarise(occurence = n()) %>%
  arrange (desc(occurence))


attorney_trust_account <- md_attorney_sanctions %>%
  filter (str_detect(text, "attorney trust account" ))
```
