---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
sports_politics <- read_csv("data/sports-political-donations.csv") %>%
clean_names()

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**:“Of the 160 owners and commissioners in ESPN’s database, 74 contributed a majority of their publicly searchable funds to Republican campaigns, with 48 donating a majority to Democrats and the rest sending most of their money to causes that aren’t fully aligned with either party.”

* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
sports_politics %>%
  group_by(party,owner) %>%
summarise (count_party=n())%>%
 pivot_wider(names_from=party, values_from=count_party)

# Display results of code below this codeblock
```

### Sentence 2

* **Sentence text**: But while that overall summary is interesting, a more pronounced signal begins to emerge when we look more closely at this group’s larger political donations. There are 40 owners who have donated at least 100,000 to Republican causes over the six-year period, compared to 23 owners topping $100,000 to Democrats, helping lead to these splits
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
sports_politics <- sports_politics_filtered 
sports_politics$amount <- (gsub(",", "", sports_politics$amount))
sports_politics$amount <- gsub("\\$", "", sports_politics$amount)
sports_politics$amount <- as.numeric(sports_politics$amount)
```
```{r}
sports_politics %>%
  group_by(owner)%>%
  filter(party=="Republican", amount>=100000)
```
```{r}

```

### Sentence 3

* **Sentence text**: “Overall, 51.7 percent of all contributions by WNBA owners went to Democrats, with 42.3 percent going to Republicans. And of the $1.3 million contributed by WNBA owners to Republicans, 65.5 percent comes from Kelly Loeffler, the Republican from Georgia who is co-owner of the Atlanta Dream.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
sports_politics %>% 
  group_by(league,party)%>% 
  filter(league=="WNBA")

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-