---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
sports_politics <- read_csv("data/sports-political-donations.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: But while that overall summary is interesting, a more pronounced signal begins to emerge when we look more closely at this group’s larger political donations. There are 40 owners who have donated at least $100,000 to Republican causes over the six-year period, compared to 23 owners topping $100,000 to Democrats, helping lead to these splits:” 
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
sports_politics_filtered <- sports_politics %>%
  
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: 
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
sports_politics_filtered <- sports_politics %>%
  
print(sapply(sports_politics_filtered, class)) 

#Convert column1 column type to numeric
sports_politics_filtered$Amount = as.numeric(as.character(sports_politics_filtered$Amount)) 

#Return the datatype of each column
print(sapply(sports_politics_filtered, class)) 

#Return the datatype of each column by converting column to numeric with transform()
print(sapply(transform(sports_politics_filtered,Amount = as.numeric(Amount)), class))
  
# Display results of code below this codeblock

glimpse (sports_politics_filtered)

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-