---
title: "lab_09.Rmd"
author: "derek willis"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Datawrapper account
* Data in CSV files in the `data` folder

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# load libraries here
```

## Questions

**Q1.** You want to explore the differences between the electoral performances of Dan Cox in 2022 and Larry Hogan in 2018. Except a map is probably not the best way to show that. Let's make a chart instead, and the type is up to you, but the goal is to clearly show the differences in total votes between the two.

You MUST do the following things when answering this question:
* Load the md_county_gov.csv and md_gov_county_18.csv files from the data folder
* Join those dataframes, specifying the join column. 
* Add columns to generate percentages for Cox and Hogan.
* Export that joined dataframe to a CSV file using `write_csv`.
* Using that exported CSV, make a chart with Datawrapper. What kind of chart is up to you, so experiment with a few kinds of charts to see what answers the question above in the clearest way. Your chart MUST:
  * Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.
  * Easily readable labels and values
* Publish your chart
* Describe your thought process below - why you chose to make the chart you did - and add the URL of your chart.

**A1.For my visualization I wanted to highlight the fact that each county had a higher percentage for Hogan than for Cox so I made a comparative dot plot which chrted each data point for each county and compared the two. I feel like this was a very easy way to see the difference between the two percentage data sets in that all the Cox data points fall to the left (lower numbers) and all the Hogan data points fall to the right (higher numbers):  https://www.datawrapper.de/_/XlD3Y/** 

```{r}
library(tidyverse)
library(janitor)

md_gov_county <- read_csv("data/md_gov_county.csv")
md_gov_county_18 <- read_csv("data/md_gov_county_18.csv")
merge_md_gov_county<- md_gov_county %>% left_join(md_gov_county_18, by=c('GEOID'))
merge_md_gov_county <- merge_md_gov_county %>% 
  mutate(total_votes_2022 = Cox + Moore + Lashar + Harding + Wallace + Sefcik + `Write-ins.x`) %>%
  mutate(pct_cox = Cox/total_votes_2022 * 100) %>%
  mutate(total_votes_2018 = Hogan + Jealous + Quinn + Schlakman + `Write-ins.y`) %>%
  mutate(pct_hogan = Hogan/total_votes_2018 * 100)
write_csv(merge_md_gov_county, "data/merge_md_gov_county.csv")
```

**Q2.** You're interested in seeing how the methods of voting have changed in Maryland since 2018. Using the `md_votes_by_type.csv` that has data from the top of the ticket (governor in 2018 and 2022, president in 2020), make a chart showing the percentage of votes cast in each category - early voting, election day and mail - over the three years. You'll need to make those percentages in R and then export the CSV using `write_csv`, then make the chart in Datawrapper. Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.

HINT: In Datawrapper's "Check and Describe" tab, you can click on a column to get the option of "hiding" it in any visualization.

What do you think the story here is? Put the answer below along with the URL of your chart.

**A2.I believe the story here is how the Covid-19 pandemic affected the way that people voted in 2020. There was a massive shift from people voting in-person on election day to people voting by mail instead, which I believe could have been influenced by voters wanting to stay as phyiscally distant as possible. https://www.datawrapper.de/_/ubjiT/** 

```{r}
md_votes_by_type <- read_csv("data/md_votes_by_type.csv")

```


**Q3.** You want to know more about the geography of 911 overdose calls in Montgomery County, Maryland, in 2022. To do that, you'll make a choropleth map in Datawrapper. You will need to:

*Read the file into R and create zip code totals and then export that CSV into your map
* Search for Maryland zip codes (use the 2020 data) as the base map
* Upload the CSV file
* Change the tooltip to show both the number and the percentage (you can put them side by side or put one under the other). Make sure to properly format both numbers to make them readable.
* Have a title, description and source that tells readers about the data. The title and description should highlight the most newsworthy or important aspect.

What do you think the story here is, and what would you want to know more about? Put the answer below along with the URL of your map.

**A3.The story here is that Germantown had such a high number of calls in comparison to the other zipcodes, I definitely would want to look more into why that seems to be the case and where specifically in the Germantown zip code that these 911 calls are coming from. https://www.datawrapper.de/_/AFck8/** 
```{r}
montgomery_2022_overdoses <- read_csv("data/montgomery_2022_overdoses.csv")
montgomery_2022_calls <- montgomery_2022_overdoses %>% 
group_by(zip)%>% 
  summarize(total_calls = n())
write_csv(montgomery_2022_calls, "data/montgomery_2022_calls.csv")
```
